---
title: "linear model from Web"
author: "Shaun Ikeuchi"
date: "2/28/24"
output:
   html_document:
         toc: true
         toc_depth: 5
        
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Read about the data from the website where it is staged.  Then form three regression models; Model1 is SIMS~ARM, Model2 is SIMS~GRIP, and Model3 is SIMS~ARM+GRIP. For each model find a 95% prediction interval of SIMS given a value of 94  for GRIP and 88 for ARM. Compare Model1 with Model3 using anova. Write it up in a markdown document, push the project up to your github account and submit it back to canvas as link. 

 


```{r include=FALSE}
require(tidyverse)
require(tigerstats)
require(rgl)
require(knitr)

```

```{r}
data <- read.csv(file="https://onlinestatbook.com/case_studies_rvls/physical_strength/data.txt",sep="",header=TRUE)  

```
# Introduction
We will produce three equation models to get 
 
1. sim score based on arm strength 
 
2. sim score based on grip strength 
 
3. sim score based on both arm and grip strength 
 

For each model we will predict sim given a value of 94 for GRIP and 88 for ARM.
 
# Methods
 
We have numerical response (sim score) and numerical explanatory variables (arm and group strength) so we use the numerical~numerical helper file.
 
## Model 1 SIM~ARMS

### scatterplot
```{r}
plot(SIMS~ARM, data=data)
```
Based on the scatterplot, there is a positive association between SIMS and ARM. If this was a perfect linear relationship, all points would be on a straight line going up.
### Numerical results
```{r}
cor(SIMS~ARM, data=data)
```
 
The correlation between SIM and arm strength is 0.680073.
### Inferential  (Build model.1)

```{r}
model.1 <- lm(SIMS~ARM,data=data)
summary.lm(model.1)
```
The p-value for the coefficient of arm is 0.00000000000000002.
#### Predict at target point
```{r}
SIMSpredict = predict.lm(model.1, newdata = data.frame(ARM=88, GRIP=94),interval="confidence")
print(SIMSpredict)
```
 
The prediction for AMR=88 and GRIP=94 is 0.71.
#### scatterplot with model fit
```{r}
plot(SIMS~ARM,data=data)
abline(model.1, col="red")
``` 

 This is a scatter plot with the line of best fit. The ARM strength at 88 has a 95 percent confidence interval between 0.49 and 0.92.
## Model 2 SIM~GRIP

### Now add in scatterplot
```{r}
plot(SIMS~GRIP, data=data)
```
 
Based on the scatterplot, there is a positive association between SIMS and GRIP. If this was a perfect linear relationship, all points would be on a straight line going up.
### Numerical results 
```{r}
cor(SIMS~GRIP, data=data)
```
 
The correlation between SIM and arm strength is 0.69398458.
### Inferential  (Build model.1)

```{r}
model.2 <- lm(SIMS~GRIP,data=data)
summary.lm(model.2)
```
 
The p-value for the coefficient of arm is 0.00000000000000002.
#### predict model.2 at target point
 
```{r}
SIMSpredict = predict.lm(model.2, newdata = data.frame(ARM=88, GRIP=94),interval="confidence")
print(SIMSpredict)
```
 
The prediction for ARM=88 and GRIP=94 is -0.535143.
#### now add the model fit to our plot for model.2
  
```{r}
plot(SIMS~GRIP,data=data)
abline(model.2, col="red")
``` 
 
 This is a scatter plot with the line of best fit. The GRIP strength at 94 has a 95 percent confidence interval between -0.7924943 -0.2798144.

## Model 3 SIM~ARM+GRIP

### Numerical results (cor)
 
```{r}
m = data.frame(data['GRIP'], data['ARM'], data['SIMS'])
cor(m)
```
 
The table shows that SIMS correlates with GRIP in a positive manner(0.64) and SIMS correlates with ARM also in a positive manner (0.69). This shows that these variables are useful in predicting SIMS.
  
```{r}
plot(data$GRIP, data$ARM)

```
### Inferential  (Build 2-dimentional model.3)
```{r}
model.3 = lm(SIMS~GRIP+ARM,data=data)
print(model.3)
newPoint = data.frame(GRIP=94, ARM=88)
predict.lm(model.3, newPoint, interval='prediction')
```
 
The 95% confidence interval in SIMS is between -2.13 and 2.43.  Since this interval contains 0, we do not have a very good prediction for SIMS.

#### predict model.3 at target point
  
```{r}
SIMSpredict = predict.lm(model.3, newdata = data.frame(ARM=88, GRIP=94))
print(SIMSpredict)
```
## Comparing nested models ANOVA Test

### Model.1 vs Model.3

```{r}
anova(model.1, model.3)

```
 
The p-value is 0.000004994, this is less than 0.05, so we can conclude that model.3 is better than model.1.



### Model.2 vs Model.3

## Informally compare Model.1 with model.2
```{r}
anova(model.2, model.3)

```


The p-value is 0.000000004995, this is less than 0.05, so we can conclude that model.3 is better than model.2.
